How do amateur photometric observations of binary star transits compare in precision and reliability to data collected by professional astronomers?

### 2024-12-25 paper.tex
docs: writting
Signed-off-by: Thomas Mauran <thomasmauran@yahoo.com>

% Introduction
\section{Introduction}
In recent years, the field of astronomy has witnessed a growing collaboration between amateur and professional astronomers. The advent of high-quality, low-cost equipment has enabled amateur astronomers to perform sophisticated observations, including photometric studies of binary star transits. This paper focuses on photometric observations of the U Cepheihei (U Cephei) binary star system conducted by the author using personal equipment from a Bortle 7 location. The study aims to examine the precision and reliability of these amateur observations in comparison to professional datasets, exploring the potential for collaboration and the challenges inherent in combining data from diverse sources.

>>>>

% skip to the next page
\newpage

% Table of Contents
\tableofcontents

% Introduction
\section{Introduction}
In recent years, the field of astronomy has witnessed a growing collaboration between amateur 
and professional astronomers. The advent of high-quality, low-cost equipment has enabled 
amateur astronomers to perform sophisticated observations, including photometric studies of 
binary star transits. This paper focuses on photometric observations of the U Cephei 
(U Cephei) binary star system conducted by the author using personal equipment from a 
Bortle 7 location. The study aims to examine the precision and reliability of these amateur 
observations in comparison to professional datasets, exploring the potential for collaboration
and the challenges inherent in combining data from diverse sources.

% Problematic
\subsection{Problematic}

% State of the Art
\section{State of the Art}
The field of photometric observation has witnessed significant advancements due to improvements in both professional and amateur technologies.

\section{What are transits?}

Transits are a phenomenon that occurs when a celestial body passes in front of another, partially or totally blocking its light.
Those events are very interesting for astronomers as they can provide a lot of information about the objects involved, such as their size, mass, and even atmosphere.

The first exoplanet was discovered using the transit method, and it's still one of the most used methods to detect exoplanets.
One of the way to know if a star is part of a binary system is to observe it and see if it's light intensity varies over time, which can be a sign of a transit event.

\section{On the U Cephei system}

U Cephei is a very interesting binary star system and a great pick to try catching transit eclipses.

\subsection{Why U Cephei?}

There are multiple reasons why U Cephei is a great candidate for me in this study.

\subsubsection{Period}

First of all U Cephei period is 2.493 days, which is quite short and allows for multiple chances to observe it on a short period of time. 
Even though this period seems short since it's not exactly 3 days but 2 and a half it means that one observation out of 2 will be at night and the other during the day.
Making observation nights possible every 5 days.
Considering the fact weather conditions are critical for this field, managing to have a night with good seeing condition can also be a challenge, so having multiple chances to observe the same system is a great advantage.

\subsubsection{Eclipse duration}

Secondly, this system eclipse is short, lasting only a couple of hours it's possible in one night to observe the complete eclipse, which is not the case for all binary systems.

\subsubsection{Magnitude variation}

Stars magnitude define how bright they are. This scale of magnitude is logarithmic, meaning that a difference of 1 magnitude is a difference of 2.512 times in brightness and a difference of 2 magnitudes is a difference of \(\mathbf{2.512^2}\) times in brightness.
the scale is inverted, meaning that the lower the magnitude, the brighter the star is.

The magnitude of U Cephei varies from 6.7 to 9.2. Which is actually a huge variation in brightness. If we wanna have a comparaison in percentage we can do the following calculation:

We can right the magnitude variation as a ratio of brightness variation:

\begin{equation}
     2.512^{m_2-m_1} = \frac{F_1}{F_2}
\end{equation}

\begin{equation}
    2.512^{9.2-6.7} \approx 10
\end{equation}

In other words, during the transit, U Cephei will be 10 times less bright than when it's not in transit. This is a huge variation and makes it easier to detect the transit.

\subsubsection{Data availability}

Finally, U Cephei is a well-known system, with a lot of data available from professional observatories, which makes it a great candidate for me as an amateur astronomer to try to catch some transits and compare my data to the professional ones.
